Weakly Nonlinear MRI: Notes

### 2014-03-15 notes/mri_notes.tex
nonlinear vector
more nonlinear expansion

$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - J(A, \nabla^2 A)\\
J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
-J(A, \Psi) \\
J(\Psi, B_{1y} - J(A, u_{1y}) \\
\end{matrix} \right]$ \\

@@

$\partial_t\mathcal{D}\mathbf{V} \, + \, \epsilon^2 \partial_T \mathcal{D}\mathbf{V} \, + \, \epsilon^2\partial_Z^2\partial_t\mathcal{D^*}\mathbf{V} \, + \, 2\epsilon\partial_z\partial_Z\partial_t\mathcal{D^*}\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$

>>>>

$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - J(A, \nabla^2 A)\\
J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
-J(A, \Psi) \\
J(\Psi, B_{1y}) - J(A, u_{1y}) \\
\end{matrix} \right]$ \\

@@

$\mathcal{D}\partial_t\mathbf{V} \, + \, \epsilon^2  \mathcal{D}\partial_T\mathbf{V} \, + \, \epsilon^2\mathcal{D^*}\partial_Z^2\partial_t\mathbf{V} \, + \, 2\epsilon\mathcal{D^*}\partial_z\partial_Z\partial_t\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$ \\

What about the vector of nonlinear terms? Let $\mathbf{N} = \left[N^{\left(\Psi\right)}, N^{\left(u_{1y}\right)}, N^{\left(A\right)}, N^{\left(B_{1y}\right)} \right]^\mathbf{T}$. We begin with the term $N^{\left(\Psi\right)} = J\left(\Psi, \nabla^2\Psi\right) - J\left(A, \nabla^2A\right)$: \\

Making the replacements: \\

$\partial_z \to \partial_z + \epsilon\partial_Z$ \\

$\nabla^2 \to \nabla^2 + \epsilon^2\partial_Z^2 + 2\epsilon\partial_z\partial_Z$ \\

We find that $J\left(\Psi, \nabla^2\Psi\right) = \partial_z\Psi\partial_x\nabla^2\Psi - \partial_x\Psi\partial_z\nabla^2\Psi$ becomes: \\

$\partial_z\Psi\partial_x\nabla^2\Psi \, - \, \partial_x\Psi\partial_z\nabla^2\Psi \, + \, \epsilon^2\partial_z\Psi\partial_x\partial_Z^2\Psi \, - \, \epsilon^2\partial_x\Psi\partial_z\partial_Z^2\Psi \, + \, 2\epsilon\partial_z\Psi\partial_x\partial_z\partial_Z\Psi \, - \, 2\epsilon\partial_x\Psi\partial_z^2\partial_Z\Psi \, + \, \epsilon\partial_Z\Psi\partial_x\nabla^2\Psi \, - \, \epsilon\partial_x\Psi\partial_Z\nabla^2\Psi \, + \, \epsilon^3\partial_Z\Psi\partial_x\partial_Z^2\Psi \, - \, \epsilon^3\partial_x\Psi\partial_Z^3\Psi \, + \, 2\epsilon^2\partial_Z\Psi\partial_x\partial_z\partial_Z\Psi \, - \, 2\epsilon^2\partial_x\Psi\partial_z\partial_Z^2\Psi$ \\

$= \, J\left(\Psi, \nabla^2\Psi\right) \, + \, \epsilon^2J\left(\Psi, \partial_Z^2\Psi\right) \, + \, 2\epsilon J\left(\Psi, \partial_z\partial_Z\Psi\right) \, + \epsilon\widetilde{J}\left(\Psi, \nabla^2\Psi\right) \, + \, \epsilon^3\widetilde{J}\left(\Psi, \partial_Z^2\Psi\right) \, + \, 2\epsilon^2\widetilde{J}\left(\Psi, \partial_z\partial_Z\Psi\right)$ \\

Where $\widetilde{J}\left(f, g\right) \equiv \partial_Z f \partial_x g - \partial_x f \partial_Z g$ \\

When we expand the fluid quantities $\mathbf{V}$ in a perturbation series $\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...$, we will introduce additional $\epsilon$ terms. Because we are truncating our expansion at $\mathcal{O}(\epsilon^3)$, we will write out the full perturbation series expansion of the nonlinear vector now. \\

Note that $J\left(a + b, c + d\right) = J(a, c) + J(a, d) + J(b, c) + J(b, d)$. \\

Thus the first term of $N^{\left(\Psi\right)}$ expands to become: \\

$J(\epsilon\Psi_1 + \epsilon^2\Psi_2, \epsilon\nabla^2 \Psi_1 + \epsilon^2 \nabla^2\Psi_2)$
